\section{Bioconductor software packages with `cancer' in package description}\label{app1}

Quoted text is the content of the Description element of the package DESCRIPTION. Names following the
text are as reported in the Authors field.

\begin{longtable}[t]{l>{\raggedright\arraybackslash}p{25em}}
\toprule
package & desc\\
\midrule
AMARETTO & "Integrating an increasing number of available multi-omics
cancer data remains one of the main challenges to improve our
understanding of cancer. One of the main challenges is using
multi-omics data for identifying novel cancer driver genes. We
have developed an algorithm, called AMARETTO, that integrates
copy number, DNA methylation and gene expression data to
identify a set of driver genes by analyzing cancer samples and
connects them to clusters of co-expressed genes, which we
define as modules. We applied AMARETTO in a pancancer setting
to identify cancer driver genes and their modules on multiple
cancer sites. AMARETTO captures modules enriched in
angiogenesis, cell cycle and EMT, and modules that accurately
predict survival and molecular subtypes. This allows AMARETTO
to identify novel cancer driver genes directing canonical
cancer pathways." --Jayendra Shinde, Celine Everaert, Shaimaa Bakr, Mohsen Nabian, Jishu Xu, Vincent Carey, Nathalie Pochet, Olivier Gevaert\\
BaalChIP & "The package offers functions to process multiple ChIP-seq
BAM files and detect allele-specific events. Computes allele
counts at individual variants (SNPs/SNVs), implements extensive
QC steps to remove problematic variants, and utilizes a
bayesian framework to identify statistically significant
allele- specific events. BaalChIP is able to account for copy
number differences between the two alleles, a known
phenotypical feature of cancer samples." --Ines de Santiago, Wei Liu, Ke Yuan, Martin O'Reilly, Chandra SR Chilamakuri, Bruce Ponder, Kerstin Meyer, Florian Markowetz\\
bioCancer & "This package is a Shiny App to visualize and analyse
interactively Multi-Assays of Cancer Genomic Data." --Karim Mezhoud\\
BiocOncoTK & "Provide a central interface to various tools for
genome-scale analysis of cancer studies." --Vince Carey\\
biodbNci & "The biodbNci library is an extension of the biodb
framework package. It provides access to biodbNci, a library
for connecting to the National Cancer Institute (USA) CACTUS
Database. It allows to retrieve entries by their accession
number, and run specific web services." --Pierrick Roger\\
\addlinespace
canceR & "The package is user friendly interface based on the cgdsr
and other modeling packages to explore, compare, and analyse
all available Cancer Data (Clinical data, Gene Mutation, Gene
Methylation, Gene Expression, Protein Phosphorylation, Copy
Number Alteration) hosted by the Computational Biology Center
at Memorial-Sloan-Kettering Cancer Center (MSKCC)." --Karim Mezhoud. Nuclear Safety \& Security Department. Nuclear Science Center of Tunisia.\\
cbaf & "This package contains functions that allow analysing and
comparing omic data across various cancers/cancer subgroups
easily. So far, it is compatible with RNA-seq, microRNA-seq,
microarray and methylation datasets that are stored on
cbioportal.org." --Arman Shahrisa, Maryam Tahmasebi Birgani\\
cBioPortalData & "The cBioPortalData R package accesses study datasets from
the cBio Cancer Genomics Portal. It accesses the data either
from the pre-packaged zip / tar files or from the API interface
that was recently implemented by the cBioPortal Data Team. The
package can provide data in either tabular format or with
MultiAssayExperiment object that uses familiar Bioconductor
data representations." --Levi Waldron, Marcel Ramos, Karim Mezhoud\\
cbpManager & "This R package provides an R Shiny application that
enables the user to generate, manage, and edit data and
metadata files suitable for the import in cBioPortal for Cancer
Genomics. Create cancer studies and edit its metadata. Upload
mutation data of a patient that will be concatenated to the
data\_mutation\_extended.txt file of the study. Create and edit
clinical patient data, sample data, and timeline data. Create
custom timeline tracks for patients." --Arsenij Ustjanzew, Federico Marini\\
ccfindR & "A collection of tools for cancer genomic data clustering
analyses, including those for single cell RNA-seq. Cell
clustering and feature gene selection analysis employ Bayesian
(and maximum likelihood) non-negative matrix factorization
(NMF) algorithm. Input data set consists of RNA count matrix,
gene, and cell bar code annotations.  Analysis outputs are
factor matrices for multiple ranks and marginal likelihood
values for each rank. The package includes utilities for
downstream analyses, including meta-gene identification,
visualization, and construction of rank-based trees for
clusters." --Jun Woo, Jinhua Wang\\
\addlinespace
cfDNAPro & "cfDNA fragments carry important features for building
cancer sample classification ML models, such as fragment size,
and fragment end motif etc. Analyzing and visualizing fragment
size metrics, as well as other biological features in a
curated, standardized, scalable, well-documented, and
reproducible way might be time intensive. This package intends
to resolve these problems and simplify the process. It offers
two sets of functions for cfDNA feature characterization and
visualization." --Haichao Wang, Hui Zhao, Elkie Chan, Christopher Smith, Tomer Kaplan, Florian Markowetz, Nitzan Rosenfeld\\
cfTools & "The cfTools R package provides methods for cell-free DNA
(cfDNA) methylation data analysis to facilitate cfDNA-based
studies. Given the methylation sequencing data of a cfDNA
sample, for each cancer marker or tissue marker, we deconvolve
the tumor-derived or tissue-specific reads from all reads
falling in the marker region. Our read-based deconvolution
algorithm exploits the pervasiveness of DNA methylation for
signal enhancement, therefore can sensitively identify a trace
amount of tumor-specific or tissue-specific cfDNA in plasma.
cfTools provides functions for (1) cancer detection:
sensitively detect tumor-derived cfDNA and estimate the
tumor-derived cfDNA fraction (tumor burden); (2) tissue
deconvolution: infer the tissue type composition and the cfDNA
fraction of multiple tissue types for a plasma cfDNA sample.
These functions can serve as foundations for more advanced
cfDNA-based studies, including cancer diagnosis and disease
monitoring." --Ran Hu, Mary Louisa Stackpole, Shuo Li, Xianghong Jasmine Zhou, Wenyuan Li\\
CIMICE & "CIMICE is a tool in the field of tumor phylogenetics and
its goal is to build a Markov Chain (called Cancer Progression
Markov Chain, CPMC) in order to model tumor subtypes evolution.
The input of CIMICE is a Mutational Matrix, so a boolean matrix
representing altered genes in a collection of samples. These
samples are assumed to be obtained with single-cell DNA
analysis techniques and the tool is specifically written to use
the peculiarities of this data for the CMPC construction." --Nicolò Rossi\\
compSPOT & "Clonal cell groups share common mutations within cancer,
precancer, and even clinically normal appearing tissues. The
frequency and location of these mutations may predict prognosis
and cancer risk. It has also been well established that certain
genomic regions have increased sensitivity to acquiring
mutations. Mutation-sensitive genomic regions may therefore
serve as markers for predicting cancer risk. This package
contains multiple functions to establish significantly mutated
hotspots, compare hotspot mutation burden between samples, and
perform exploratory data analysis of the correlation between
hotspot mutation burden and personal risk factors for cancer,
such as age, gender, and history of carcinogen exposure. This
package allows users to identify robust genomic markers to help
establish cancer risk." --Sydney Grant, Ella Sampson, Rhea Rodrigues, Gyorgy Paragh\\
consensusOV & "This package implements four major subtype classifiers for
high-grade serous (HGS) ovarian cancer as described by Helland
et al. (PLoS One, 2011), Bentink et al. (PLoS One, 2012),
Verhaak et al. (J Clin Invest, 2013), and Konecny et al. (J
Natl Cancer Inst, 2014). In addition, the package implements a
consensus classifier, which consolidates and improves on the
robustness of the proposed subtype classifiers, thereby
providing reliable stratification of patients with HGS ovarian
tumors of clearly defined subtype." --Gregory M Chen, Lavanya Kannan, Ludwig Geistlinger, Victor Kofia, Levi Waldron, Benjamin Haibe-Kains\\
\addlinespace
copa & "COPA is a method to find genes that undergo recurrent
fusion in a given cancer type by finding pairs of genes that
have mutually exclusive outlier profiles." --James W. MacDonald\\
dce & "Compute differential causal effects (dce) on (biological)
networks. Given observational samples from a control experiment
and non-control (e.g., cancer) for two genes A and B, we can
compute differential causal effects with a (generalized) linear
regression. If the causal effect of gene A on gene B in the
control samples is different from the causal effect in the
non-control samples the dce will differ from zero. We
regularize the dce computation by the inclusion of prior
network information from pathway databases such as KEGG." --Kim Philipp Jablonski, Martin Pirkl\\
DepInfeR & "DepInfeR integrates two experimentally accessible input
data matrices: the drug sensitivity profiles of cancer cell
lines or primary tumors ex-vivo (X), and the drug affinities of
a set of proteins (Y), to infer a matrix of molecular protein
dependencies of the cancers (ß). DepInfeR deconvolutes the
protein inhibition effect on the viability phenotype by using
regularized multivariate linear regression. It assigns a
“dependence coefficient” to each protein and each sample, and
therefore could be used to gain a causal and accurate
understanding of functional consequences of genomic aberrations
in a heterogeneous disease, as well as to guide the choice of
pharmacological intervention for a specific cancer type,
sub-type, or an individual patient. For more information,
please read out preprint on bioRxiv:
https://doi.org/10.1101/2022.01.11.475864." --Junyan Lu, Alina Batzilla\\
DriverNet & "DriverNet is a package to predict functional important
driver genes in cancer by integrating genome data (mutation and
copy number variation data) and transcriptome data (gene
expression data). The different kinds of data are combined by
an influence graph, which is a gene-gene interaction network
deduced from pathway data. A greedy algorithm is used to find
the possible driver genes, which may mutated in a larger number
of patients and these mutations will push the gene expression
values of the connected genes to some extreme values." --Ali Bashashati, Reza Haffari, Jiarui Ding, Gavin Ha, Kenneth Liu, Jamie Rosner, Sohrab Shah\\
easier & "This package provides a workflow for the use of EaSIeR
tool, developed to assess patients' likelihood to respond to
ICB therapies providing just the patients' RNA-seq data as
input. We integrate RNA-seq data with different types of prior
knowledge to extract quantitative descriptors of the tumor
microenvironment from several points of view, including
composition of the immune repertoire, and activity of intra-
and extra-cellular communications. Then, we use multi-task
machine learning trained in TCGA data to identify how these
descriptors can simultaneously predict several state-of-the-art
hallmarks of anti-cancer immune response. In this way we derive
cancer-specific models and identify cancer-specific systems
biomarkers of immune response. These biomarkers have been
experimentally validated in the literature and the performance
of EaSIeR predictions has been validated using independent
datasets form four different cancer types with patients treated
with anti-PD1 or anti-PDL1 therapy." --Oscar Lapuente-Santana, Federico Marini, Arsenij Ustjanzew, Francesca Finotello, Federica Eduati\\
\addlinespace
GDCRNATools & "This is an easy-to-use package for downloading,
organizing, and integrative analyzing RNA expression data in
GDC with an emphasis on deciphering the lncRNA-mRNA related
ceRNA regulatory network in cancer. Three databases of
lncRNA-miRNA interactions including spongeScan, starBase, and
miRcode, as well as three databases of mRNA-miRNA interactions
including miRTarBase, starBase, and miRcode are incorporated
into the package for ceRNAs network construction. limma, edgeR,
and DESeq2 can be used to identify differentially expressed
genes/miRNAs. Functional enrichment analyses including GO,
KEGG, and DO can be performed based on the clusterProfiler and
DO packages. Both univariate CoxPH and KM survival analyses of
multiple genes can be implemented in the package. Besides some
routine visualization functions such as volcano plot, bar plot,
and KM plot, a few simply shiny apps are developed to
facilitate visualization of results on a local webpage." --Ruidong Li, Han Qu, Shibo Wang, Julong Wei, Le Zhang, Renyuan Ma, Jianming Lu, Jianguo Zhu, Wei-De Zhong, Zhenyu Jia\\
genefu & "This package contains functions implementing various tasks
usually required by gene expression analysis, especially in
breast cancer studies: gene mapping between different
microarray platforms, identification of molecular subtypes,
implementation of published gene signatures, gene selection,
and survival analysis." --Deena M.A. Gendoo, Natchar Ratanasirigulchai, Markus S. Schroeder, Laia Pare, Joel S Parker, Aleix Prat, Nikta Feizi, Christopher Eeles, Benjamin Haibe-Kains\\
GeoTcgaData & "Gene Expression Omnibus(GEO) and The Cancer Genome Atlas
(TCGA) provide us with a wealth of data, such as RNA-seq, DNA
Methylation, SNP and Copy number variation data. It's easy to
download data from TCGA using the gdc tool, but processing
these data into a format suitable for bioinformatics analysis
requires more work. This R package was developed to handle
these data." --Erqiang Hu\\
INDEED & "An R package for integrated differential expression and
differential network analysis based on omic data for cancer
biomarker discovery. Both correlation and partial correlation
can be used to generate differential network to aid the
traditional differential expression analysis to identify
changes between biomolecules on both their expression and
pairwise association levels. A detailed description of the
methodology has been published in Methods journal (PMID:
27592383). An interactive visualization feature allows for the
exploration and selection of candidate biomarkers." --Yiming Zuo, Kian Ghaffari, Zhenzhi Li\\
iPath & "iPath is the Bioconductor package used for calculating
personalized pathway score and test the association with
survival outcomes. Abundant single-gene biomarkers have been
identified and used in the clinics. However, hundreds of
oncogenes or tumor-suppressor genes are involved during the
process of tumorigenesis. We believe individual-level
expression patterns of pre-defined pathways or gene sets are
better biomarkers than single genes. In this study, we devised
a computational method named iPath to identify prognostic
biomarker pathways, one sample at a time. To test its utility,
we conducted a pan-cancer analysis across 14 cancer types from
The Cancer Genome Atlas and demonstrated that iPath is capable
of identifying highly predictive biomarkers for clinical
outcomes, including overall survival, tumor subtypes, and tumor
stage classifications. We found that pathway-based biomarkers
are more robust and effective than single genes." --Kenong Su, Zhaohui Qin\\
\addlinespace
LACE & "LACE is an algorithmic framework that processes
single-cell somatic mutation profiles from cancer samples
collected at different time points and in distinct experimental
settings, to produce longitudinal models of cancer evolution.
The approach solves a Boolean Matrix Factorization problem with
phylogenetic constraints, by maximizing a weighed likelihood
function computed on multiple time points." --Daniele Ramazzotti, Fabrizio Angaroni, Davide Maspero, Alex Graudenzi, Luca De Sano, Gianluca Ascolani\\
macat & "This library contains functions to investigate links
between differential gene expression and the chromosomal
localization of the genes. MACAT is motivated by the common
observation of phenomena involving large chromosomal regions in
tumor cells. MACAT is the implementation of a statistical
approach for identifying significantly differentially expressed
chromosome regions. The functions have been tested on a
publicly available data set about acute lymphoblastic leukemia
(Yeoh et al.Cancer Cell 2002), which is provided in the library
'stjudem'." --Benjamin Georgi, Matthias Heinig, Stefan Roepcke, Sebastian Schmeier, Joern Toedling\\
maftools & "Analyze and visualize Mutation Annotation Format (MAF)
files from large scale sequencing studies. This package
provides various functions to perform most commonly used
analyses in cancer genomics and to create feature rich
customizable visualzations with minimal effort." --Anand Mayakonda\\
mastR & "mastR is an R package designed for automated screening of
signatures of interest for specific research questions. The
package is developed for generating refined lists of signature
genes from multiple group comparisons based on the results from
edgeR and limma differential expression (DE) analysis workflow.
It also takes into account the background noise of
tissue-specificity, which is often ignored by other marker
generation tools. This package is particularly useful for the
identification of group markers in various biological and
medical applications, including cancer research and
developmental biology." --Jinjin Chen, Ahmed Mohamed, Chin Wee Tan\\
MethylMix & "MethylMix is an algorithm implemented to identify hyper
and hypomethylated genes for a disease. MethylMix is based on a
beta mixture model to identify methylation states and compares
them with the normal DNA methylation state. MethylMix uses a
novel statistic, the Differential Methylation value or DM-value
defined as the difference of a methylation state with the
normal methylation state. Finally, matched gene expression data
is used to identify, besides differential, functional
methylation states by focusing on methylation changes that
effect gene expression. References: Gevaert 0. MethylMix: an R
package for identifying DNA methylation-driven genes.
Bioinformatics (Oxford, England). 2015;31(11):1839-41.
doi:10.1093/bioinformatics/btv020. Gevaert O, Tibshirani R,
Plevritis SK. Pancancer analysis of DNA methylation-driven
genes using MethylMix. Genome Biology. 2015;16(1):17.
doi:10.1186/s13059-014-0579-8." --Olivier Gevaert\\
\addlinespace
Moonlight2R & "The understanding of cancer mechanism requires the
identification of genes playing a role in the development of
the pathology and the characterization of their role (notably
oncogenes and tumor suppressors). We present an updated version
of the R/bioconductor package called MoonlightR, namely
Moonlight2R, which returns a list of candidate driver genes for
specific cancer types on the basis of omics data integration.
The Moonlight framework contains a primary layer where gene
expression data and information about biological processes are
integrated to predict genes called oncogenic mediators, divided
into putative tumor suppressors and putative oncogenes. This is
done through functional enrichment analyses, gene regulatory
networks and upstream regulator analyses to score the
importance of well-known biological processes with respect to
the studied cancer type. By evaluating the effect of the
oncogenic mediators on biological processes or through random
forests, the primary layer predicts two putative roles for the
oncogenic mediators: i) tumor suppressor genes (TSGs) and ii)
oncogenes (OCGs). As gene expression data alone is not enough
to explain the deregulation of the genes, a second layer of
evidence is needed. We have automated the integration of a
secondary mutational layer through new functionalities in
Moonlight2R. These functionalities analyze mutations in the
cancer cohort and classifies these into driver and passenger
mutations using the driver mutation prediction tool,
CScape-somatic. Those oncogenic mediators with at least one
driver mutation are retained as the driver genes. As a
consequence, this methodology does not only identify genes
playing a dual role (e.g. TSG in one cancer type and OCG in
another) but also helps in elucidating the biological processes
underlying their specific roles. In particular, Moonlight2R can
be used to discover OCGs and TSGs in the same cancer type. This
may for instance help in answering the question whether some
genes change role between early stages (I, II) and late stages
(III, IV). In the future, this analysis could be useful to
determine the causes of different resistances to
chemotherapeutic treatments." --Mona Nourbakhsh, Astrid Saksager, Nikola Tom, Xi Steven Chen, Antonio Colaprico, Catharina Olsen, Matteo Tiberti, Elena Papaleo\\
MoonlightR & "Motivation: The understanding of cancer mechanism requires
the identification of genes playing a role in the development
of the pathology and the characterization of their role
(notably oncogenes and tumor suppressors). Results: We present
an R/bioconductor package called MoonlightR which returns a
list of candidate driver genes for specific cancer types on the
basis of TCGA expression data. The method first infers gene
regulatory networks and then carries out a functional
enrichment analysis (FEA) (implementing an upstream regulator
analysis, URA) to score the importance of well-known biological
processes with respect to the studied cancer type. Eventually,
by means of random forests, MoonlightR predicts two specific
roles for the candidate driver genes: i) tumor suppressor genes
(TSGs) and ii) oncogenes (OCGs). As a consequence, this
methodology does not only identify genes playing a dual role
(e.g. TSG in one cancer type and OCG in another) but also helps
in elucidating the biological processes underlying their
specific roles. In particular, MoonlightR can be used to
discover OCGs and TSGs in the same cancer type. This may help
in answering the question whether some genes change role
between early stages (I, II) and late stages (III, IV) in
breast cancer. In the future, this analysis could be useful to
determine the causes of different resistances to
chemotherapeutic treatments." --Antonio Colaprico, Catharina Olsen, Matthew H. Bailey, Gabriel J. Odom, Thilde Terkelsen, Mona Nourbakhsh, Astrid Saksager, Tiago C. Silva, André V. Olsen, Laura Cantini, Andrei Zinovyev, Emmanuel Barillot, Houtan Noushmehr, Gloria Bertoli, Isabella Castiglioni, Claudia Cava, Gianluca Bontempi, Xi Steven Chen, Elena Papaleo, Matteo Tiberti\\
NoRCE & "While some non-coding RNAs (ncRNAs) are assigned critical
regulatory roles, most remain functionally uncharacterized.
This presents a challenge whenever an interesting set of ncRNAs
needs to be analyzed in a functional context. Transcripts
located close-by on the genome are often regulated together.
This genomic proximity on the sequence can hint to a functional
association. We present a tool, NoRCE, that performs cis
enrichment analysis for a given set of ncRNAs. Enrichment is
carried out using the functional annotations of the coding
genes located proximal to the input ncRNAs. Other biologically
relevant information such as topologically associating domain
(TAD) boundaries, co-expression patterns, and miRNA target
prediction information can be incorporated to conduct a richer
enrichment analysis. To this end, NoRCE includes several
relevant datasets as part of its data repository, including
cell-line specific TAD boundaries, functional gene sets, and
expression data for coding \& ncRNAs specific to cancer.
Additionally, the users can utilize custom data files in their
investigation. Enrichment results can be retrieved in a tabular
format or visualized in several different ways. NoRCE is
currently available for the following species: human, mouse,
rat, zebrafish, fruit fly, worm, and yeast." --Gulden Olgun\\
octad & "OCTAD provides a platform for virtually screening
compounds targeting precise cancer patient groups. The
essential idea is to identify drugs that reverse the gene
expression signature of disease by tamping down over-expressed
genes and stimulating weakly expressed ones. The package offers
deep-learning based reference tissue selection, disease gene
expression signature creation, pathway enrichment analysis,
drug reversal potency scoring, cancer cell line selection, drug
enrichment analysis and in silico hit validation. It currently
covers \textasciitilde{}20,000 patient tissue samples covering 50 cancer types,
and expression profiles for \textasciitilde{}12,000 distinct compounds." --E. Chekalin, S. Paithankar, B. Zeng, B. Glicksberg, P. Newbury, J. Xing, K. Liu, A. Wen, D. Joseph, B. Chen\\
oncoscanR & "The software uses the copy number segments from a text
file and identifies all chromosome arms that are globally
altered and computes various genome-wide scores. The following
HRD scores (characteristic of BRCA-mutated cancers) are
included: LST, HR-LOH, nLST and gLOH. the package is tailored
for the ThermoFisher Oncoscan assay analyzed with their
Chromosome Alteration Suite (ChAS) but can be adapted to any
input." --Yann Christinat, Geneva University Hospitals\\
\addlinespace
OncoScore & "OncoScore is a tool to measure the association of genes to
cancer based on citation frequencies in biomedical literature.
The score is evaluated from PubMed literature by dynamically
updatable web queries." --Luca De Sano, Carlo Gambacorti Passerini, Rocco Piazza, Daniele Ramazzotti, Roberta Spinelli\\
OncoSimulR & "Functions for forward population genetic simulation in
asexual populations, with special focus on cancer progression.
Fitness can be an arbitrary function of genetic interactions
between multiple genes or modules of genes, including
epistasis, order restrictions in mutation accumulation, and
order effects. Fitness (including just birth, just death, or
both birth and death) can also be a function of the relative
and absolute frequencies of other genotypes (i.e.,
frequency-dependent fitness). Mutation rates can differ between
genes, and we can include mutator/antimutator genes (to model
mutator phenotypes). Simulating multi-species scenarios and
therapeutic interventions, including adaptive therapy, is also
possible. Simulations use continuous-time models and can
include driver and passenger genes and modules. Also included
are functions for: simulating random DAGs of the type found in
Oncogenetic Trees, Conjunctive Bayesian Networks, and other
cancer progression models; plotting and sampling from single or
multiple realizations of the simulations, including single-cell
sampling; plotting the parent-child relationships of the
clones; generating random fitness landscapes (Rough Mount Fuji,
House of Cards, additive, NK, Ising, and Eggbox models) and
plotting them." --Ramon Diaz-Uriarte, Sergio Sanchez-Carrillo, Juan Antonio Miguel Gonzalez, Alberto Gonzalez Klein, Javier Mu\textbackslash{}\textasciitilde{}noz Haro, Javier Lopez Cano, Niklas Endres, Mark Taylor, Arash Partow, Sophie Brouillet, Sebastian Matuszewski, Harry Annoni, Luca Ferretti, Guillaume Achaz, Tymoteusz Wolodzko, Guillermo Gorines Cordero, Ivan Lorca Alonso, Francisco Mu\textbackslash{}\textasciitilde{}noz Lopez, David Roncero Moro\textbackslash{}\textasciitilde{}no, Alvaro Quevedo, Pablo Perez, Cristina Devesa, Alejandro Herrador, Holger Froehlich, Florian Markowetz, Achim Tresch, Theresa Niederberger, Christian Bender, Matthias Maneck, Claudio Lottaz, Tim Beissbarth, Sara Dorado Alfaro, Miguel Hernandez del Valle, Alvaro Huertas Garcia, Diego Ma\textbackslash{}\textasciitilde{}nanes Cayero, Alejandro Martin Mu\textbackslash{}\textasciitilde{}noz, Marta Couce Iglesias, Silvia Garcia Cobos, Carlos Madariaga Aramendi, Ana Rodriguez Ronchel, Lucia Sanchez Garcia, Yolanda Benitez Quesada, Asier Fernandez Pato, Esperanza Lopez Lopez, Alberto Manuel Parra Perez, Jorge Garcia Calleja, Ana del Ramo Galian, Alejandro de los Reyes Benitez, Guillermo Garcia Hoyos, Rosalia Palomino Cabrera, Rafael Barrero Rodriguez, Silvia Talavera Marcos\\
oppar & "The R implementation of mCOPA package published by Wang et
al. (2012). Oppar provides methods for Cancer Outlier profile
Analysis. Although initially developed to detect outlier genes
in cancer studies, methods presented in oppar can be used for
outlier profile analysis in general. In addition, tools are
provided for gene set enrichment and pathway analysis." --Chenwei Wang, Alperen Taciroglu, Stefan R Maetschke, Colleen C Nelson, Mark Ragan, Melissa Davis, Soroor Hediyeh zadeh, Momeneh Foroutan\\
ORFhunteR & "The ORFhunteR package is a R and C++ library for an
automatic determination and annotation of open reading frames
(ORF) in a large set of RNA molecules. It efficiently
implements the machine learning model based on vectorization of
nucleotide sequences and the random forest classification
algorithm. The ORFhunteR package consists of a set of functions
written in the R language in conjunction with C++. The
efficiency of the package was confirmed by the examples of the
analysis of RNA molecules from the NCBI RefSeq and Ensembl
databases. The package can be used in basic and applied
biomedical research related to the study of the transcriptome
of normal as well as altered (for example, cancer) human cells." --Vasily V. Grinev, Mikalai M. Yatskou, Victor V. Skakun, Maryna Chepeleva, Petr V. Nazarov\\
OutSplice & "An easy to use tool that can compare splicing events in
tumor and normal tissue samples using either a user generated
matrix, or data from The Cancer Genome Atlas (TCGA). This
package generates a matrix of splicing outliers that are
significantly over or underexpressed in tumors samples compared
to normal denoted by chromosome location. The package also will
calculate the splicing burden in each tumor and characterize
the types of splicing events that occur." --Joseph Bendik, Sandhya Kalavacherla, Michael Considine, Bahman Afsari, Michael F. Ochs, Joseph Califano, Daria A. Gaykalova, Elana Fertig, Theresa Guo\\
\addlinespace
pathifier & "Pathifier is an algorithm that infers pathway deregulation
scores for each tumor sample on the basis of expression data.
This score is determined, in a context-specific manner, for
every particular dataset and type of cancer that is being
investigated. The algorithm transforms gene-level information
into pathway-level information, generating a compact and
biologically relevant representation of each sample." --Yotam Drier\\
paxtoolsr & "The package provides a set of R functions for interacting
with BioPAX OWL files using Paxtools and the querying Pathway
Commons (PC) molecular interaction database. Pathway Commons is
a project by the Memorial Sloan-Kettering Cancer Center
(MSKCC), Dana-Farber Cancer Institute (DFCI), and the
University of Toronto. Pathway Commons databases include: BIND,
BioGRID, CORUM, CTD, DIP, DrugBank, HPRD, HumanCyc, IntAct,
KEGG, MirTarBase, Panther, PhosphoSitePlus, Reactome, RECON,
TRANSFAC." --Augustin Luna\\
PDATK & "Pancreatic ductal adenocarcinoma (PDA) has a relatively
poor prognosis and is one of the most lethal cancers. Molecular
classification of gene expression profiles holds the potential
to identify meaningful subtypes which can inform therapeutic
strategy in the clinical setting. The Pancreatic Cancer
Adenocarcinoma Tool-Kit (PDATK) provides an S4 class-based
interface for performing unsupervised subtype discovery,
cross-cohort meta-clustering, gene-expression-based
classification, and subsequent survival analysis to identify
prognostically useful subtypes in pancreatic cancer and beyond.
Two novel methods, Consensus Subtypes in Pancreatic Cancer
(CSPC) and Pancreatic Cancer Overall Survival Predictor (PCOSP)
are included for consensus-based meta-clustering and
overall-survival prediction, respectively. Additionally, four
published subtype classifiers and three published prognostic
gene signatures are included to allow users to easily recreate
published results, apply existing classifiers to new data, and
benchmark the relative performance of new methods. The use of
existing Bioconductor classes as input to all PDATK classes and
methods enables integration with existing Bioconductor
datasets, including the 21 pancreatic cancer patient cohorts
available in the MetaGxPancreas data package. PDATK has been
used to replicate results from Sandhu et al (2019)
[https://doi.org/10.1200/cci.18.00102] and an additional paper
is in the works using CSPC to validate subtypes from the
included published classifiers, both of which use the data
available in MetaGxPancreas. The inclusion of subtype centroids
and prognostic gene signatures from these and other
publications will enable researchers and clinicians to classify
novel patient gene expression data, allowing the direct
clinical application of the classifiers included in PDATK.
Overall, PDATK provides a rich set of tools to identify and
validate useful prognostic and molecular subtypes based on
gene-expression data, benchmark new classifiers against
existing ones, and apply discovered classifiers on novel
patient data to inform clinical decision making." --Vandana Sandhu, Heewon Seo, Christopher Eeles, Neha Rohatgi, Benjamin Haibe-Kains\\
PharmacoGx & "Contains a set of functions to perform large-scale
analysis of pharmaco-genomic data. These include the
PharmacoSet object for storing the results of pharmacogenomic
experiments, as well as a number of functions for computing
common summaries of drug-dose response and correlating them
with the molecular features in a cancer cell-line." --Petr Smirnov, Christopher Eeles, Zhaleh Safikhani, Mark Freeman, Feifei Li, Jermiah Joseph, Benjamin Haibe-Kains\\
psichomics & "Interactive R package with an intuitive Shiny-based
graphical interface for alternative splicing quantification and
integrative analyses of alternative splicing and gene
expression based on The Cancer Genome Atlas (TCGA), the
Genotype-Tissue Expression project (GTEx), Sequence Read
Archive (SRA) and user-provided data. The tool interactively
performs survival, dimensionality reduction and median- and
variance-based differential splicing and gene expression
analyses that benefit from the incorporation of clinical and
molecular sample-associated features (such as tumour stage or
survival). Interactive visual access to genomic mapping and
functional annotation of selected alternative splicing events
is also included." --Nuno Saraiva-Agostinho, Nuno Luís Barbosa-Morais, André Falcão, Lina Gallego Paez, Marie Bordone, Teresa Maia, Mariana Ferreira, Ana Carolina Leote, Bernardo de Almeida\\
\addlinespace
RadioGx & "Computational tool box for radio-genomic analysis which
integrates radio-response data, radio-biological modelling and
comprehensive cell line annotations for hundreds of cancer cell
lines. The 'RadioSet' class enables creation and manipulation
of standardized datasets including information about cancer
cells lines, radio-response assays and dose-response
indicators. Included methods allow fitting and plotting
dose-response data using established radio-biological models
along with quality control to validate results. Additional
functions related to fitting and plotting dose response curves,
quantifying statistical correlation and calculating area under
the curve (AUC) or survival fraction (SF) are included. For
more details please see the included documentation, references,
as well as: Manem, V. et al (2018) <doi:10.1101/449793>." --Venkata Manem, Petr Smirnov, Ian Smith, Meghan Lambie, Christopher Eeles, Scott Bratman, Jermiah Joseph, Benjamin Haibe-Kains\\
RAIDS & "This package implements specialized algorithms that enable
genetic ancestry inference from various cancer sequences
sources (RNA, Exome and Whole-Genome sequences). This package
also implements a simulation algorithm that generates synthetic
cancer-derived data. This code and analysis pipeline was
designed and developed for the following publication: Belleau,
P et al. Genetic Ancestry Inference from Cancer-Derived
Molecular Data across Genomic and Transcriptomic Platforms.
Cancer Res 1 January 2023; 83 (1): 49–58." --Pascal Belleau, Astrid Deschênes, David A. Tuveson, Alexander Krasnitz\\
rcellminer & "The NCI-60 cancer cell line panel has been used over the
course of several decades as an anti-cancer drug screen. This
panel was developed as part of the Developmental Therapeutics
Program (DTP, http://dtp.nci.nih.gov/) of the U.S. National
Cancer Institute (NCI). Thousands of compounds have been tested
on the NCI-60, which have been extensively characterized by
many platforms for gene and protein expression, copy number,
mutation, and others (Reinhold, et al., 2012). The purpose of
the CellMiner project (http://discover.nci.nih.gov/ cellminer)
has been to integrate data from multiple platforms used to
analyze the NCI-60 and to provide a powerful suite of tools for
exploration of NCI-60 data." --Augustin Luna, Vinodh Rajapakse, Fabricio Sousa\\
RESOLVE & "Cancer is a genetic disease caused by somatic mutations in
genes controlling key biological functions such as cellular
growth and division. Such mutations may arise both through
cell-intrinsic and exogenous processes, generating
characteristic mutational patterns over the genome named
mutational signatures. The study of mutational signatures have
become a standard component of modern genomics studies, since
it can reveal which (environmental and endogenous) mutagenic
processes are active in a tumor, and may highlight markers for
therapeutic response. Mutational signatures computational
analysis presents many pitfalls. First, the task of determining
the number of signatures is very complex and depends on
heuristics. Second, several signatures have no clear etiology,
casting doubt on them being computational artifacts rather than
due to mutagenic processes. Last, approaches for signatures
assignment are greatly influenced by the set of signatures used
for the analysis. To overcome these limitations, we developed
RESOLVE (Robust EStimation Of mutationaL signatures Via
rEgularization), a framework that allows the efficient
extraction and assignment of mutational signatures. RESOLVE
implements a novel algorithm that enables (i) the efficient
extraction, (ii) exposure estimation, and (iii) confidence
assessment during the computational inference of mutational
signatures." --Daniele Ramazzotti, Luca De Sano\\
RLassoCox & "RLassoCox is a package that implements the RLasso-Cox
model proposed by Wei Liu. The RLasso-Cox model integrates gene
interaction information into the Lasso-Cox model for accurate
survival prediction and survival biomarker discovery. It is
based on the hypothesis that topologically important genes in
the gene interaction network tend to have stable expression
changes. The RLasso-Cox model uses random walk to evaluate the
topological weight of genes, and then highlights topologically
important genes to improve the generalization ability of the
Lasso-Cox model. The RLasso-Cox model has the advantage of
identifying small gene sets with high prognostic performance on
independent datasets, which may play an important role in
identifying robust survival biomarkers for various cancer
types." --Wei Liu\\
\addlinespace
RTCGA & "The Cancer Genome Atlas (TCGA) Data Portal provides a
platform for researchers to search, download, and analyze data
sets generated by TCGA. It contains clinical information,
genomic characterization data, and high level sequence analysis
of the tumor genomes. The key is to understand genomics to
improve cancer care. RTCGA package offers download and
integration of the variety and volume of TCGA data using
patient barcode key, what enables easier data possession. This
may have an benefcial infuence on impact on development of
science and improvement of patients' treatment. Furthermore,
RTCGA package transforms TCGA data to tidy form which is
convenient to use." --Marcin Kosinski, Przemyslaw Biecek, Witold Chodor\\
RTCGAToolbox & "Managing data from large scale projects such as The Cancer
Genome Atlas (TCGA) for further analysis is an important and
time consuming step for research projects. Several efforts,
such as Firehose project, make TCGA pre-processed data publicly
available via web services and data portals but it requires
managing, downloading and preparing the data for following
steps. We developed an open source and extensible R based data
client for Firehose pre-processed data and demonstrated its use
with sample case studies. Results showed that RTCGAToolbox
could improve data management for researchers who are
interested with TCGA data. In addition, it can be integrated
with other analysis pipelines for following data analysis." --Mehmet Samur, Marcel Ramos, Ludwig Geistlinger\\
SCFA & "Subtyping via Consensus Factor Analysis (SCFA) can
efficiently remove noisy signals from consistent molecular
patterns in multi-omics data. SCFA first uses an autoencoder to
select only important features and then repeatedly performs
factor analysis to represent the data with different numbers of
factors. Using these representations, it can reliably identify
cancer subtypes and accurately predict risk scores of patients." --Duc Tran, Hung Nguyen, Tin Nguyen\\
SCOPE & "Whole genome single-cell DNA sequencing (scDNA-seq)
enables characterization of copy number profiles at the
cellular level. This circumvents the averaging effects
associated with bulk-tissue sequencing and has increased
resolution yet decreased ambiguity in deconvolving cancer
subclones and elucidating cancer evolutionary history.
ScDNA-seq data is, however, sparse, noisy, and highly variable
even within a homogeneous cell population, due to the biases
and artifacts that are introduced during the library
preparation and sequencing procedure. Here, we propose SCOPE, a
normalization and copy number estimation method for scDNA-seq
data. The distinguishing features of SCOPE include: (i)
utilization of cell-specific Gini coefficients for quality
controls and for identification of normal/diploid cells, which
are further used as negative control samples in a Poisson
latent factor model for normalization; (ii) modeling of GC
content bias using an expectation-maximization algorithm
embedded in the Poisson generalized linear models, which
accounts for the different copy number states along the genome;
(iii) a cross-sample iterative segmentation procedure to
identify breakpoints that are shared across cells from the same
genetic background." --Rujin Wang, Danyu Lin, Yuchao Jiang\\
seq.hotSPOT & "seq.hotSPOT provides a resource for designing effective
sequencing panels to help improve mutation capture efficacy for
ultradeep sequencing projects. Using SNV datasets, this package
designs custom panels for any tissue of interest and identify
the genomic regions likely to contain the most mutations.
Establishing efficient targeted sequencing panels can allow
researchers to study mutation burden in tissues at high depth
without the economic burden of whole-exome or whole-genome
sequencing. This tool was developed to make high-depth
sequencing panels to study low-frequency clonal mutations in
clinically normal and cancerous tissues." --Sydney Grant, Lei Wei, Gyorgy Paragh\\
\addlinespace
seqCNA & "Copy number analysis of high-throughput sequencing cancer
data with fast summarization, extensive filtering and improved
normalization" --David Mosen-Ansorena\\
sevenbridges & "R client and utilities for Seven Bridges platform API,
from Cancer Genomics Cloud to other Seven Bridges supported
platforms." --Phil Webster, Soner Koc, Nan Xiao, Tengfei Yin, Dusan Randjelovic, Emile Young, Velsera\\
SigCheck & "While gene signatures are frequently used to predict
phenotypes (e.g. predict prognosis of cancer patients), it it
not always clear how optimal or meaningful they are (cf David
Venet, Jacques E. Dumont, and Vincent Detours' paper "Most
Random Gene Expression Signatures Are Significantly Associated
with Breast Cancer Outcome"). Based on suggestions in that
paper, SigCheck accepts a data set (as an ExpressionSet) and a
gene signature, and compares its performance on survival and/or
classification tasks against a) random gene signatures of the
same length; b) known, related and unrelated gene signatures;
and c) permuted data and/or metadata." --Rory Stark, Justin Norden\\
signeR & "The signeR package provides an empirical Bayesian approach
to mutational signature discovery. It is designed to analyze
single nucleotide variation (SNV) counts in cancer genomes, but
can also be applied to other features as well. Functionalities
to characterize signatures or genome samples according to
exposure patterns are also provided." --Rafael Rosales, Rodrigo Drummond, Renan Valieris, Alexandre Defelicibus, Israel Tojal da Silva\\
signifinder & "signifinder is an R package for computing and exploring a
compendium of tumor signatures. It allows to compute a variety
of signatures, based on gene expression values, and return
single-sample scores. Currently, signifinder contains 46
distinct signatures collected from the literature, relating to
multiple tumors and multiple cancer processes." --Stefania Pirrotta, Enrica Calura\\
\addlinespace
supersigs & "Generate SuperSigs (supervised mutational signatures) from
single nucleotide variants in the cancer genome. Functions
included in the package allow the user to learn supervised
mutational signatures from their data and apply them to new
data. The methodology is based on the one described in Afsari
(2021, ELife)." --Albert Kuo, Yifan Zhang, Bahman Afsari, Cristian Tomasetti\\
TRONCO & "The TRONCO (TRanslational ONCOlogy) R package collects
algorithms to infer progression models via the approach of
Suppes-Bayes Causal Network, both from an ensemble of tumors
(cross-sectional samples) and within an individual patient
(multi-region or single-cell samples). The package provides
parallel implementation of algorithms that process binary
matrices where each row represents a tumor sample and each
column a single-nucleotide or a structural variant driving the
progression; a 0/1 value models the absence/presence of that
alteration in the sample. The tool can import data from plain,
MAF or GISTIC format files, and can fetch it from the
cBioPortal for cancer genomics. Functions for data manipulation
and visualization are provided, as well as functions to
import/export such data to other bioinformatics tools for, e.g,
clustering or detection of mutually exclusive alterations.
Inferred models can be visualized and tested for their
confidence via bootstrap and cross-validation. TRONCO is used
for the implementation of the Pipeline for Cancer Inference
(PICNIC)." --Marco Antoniotti, Giulio Caravagna, Luca De Sano, Alex Graudenzi, Giancarlo Mauri, Bud Mishra, Daniele Ramazzotti\\
Uniquorn & "'Uniquorn' enables users to identify cancer cell lines.
Cancer cell line misidentification and cross-contamination
reprents a significant challenge for cancer researchers. The
identification is vital and in the frame of this package based
on the locations/ loci of somatic and germline mutations/
variations. The input format is vcf/ vcf.gz and the files have
to contain a single cancer cell line sample (i.e. a single
member/genotype/gt column in the vcf file)." --Raik Otto\\
ZygosityPredictor & "The ZygosityPredictor allows to predict how many copies of
a gene are affected by small variants. In addition to the basic
calculations of the affected copy number of a variant, the
Zygosity-Predictor can integrate the influence of several
variants on a gene and ultimately make a statement if and how
many wild-type copies of the gene are left. This information
proves to be of particular use in the context of translational
medicine. For example, in cancer genomes, the
Zygosity-Predictor can address whether unmutated copies of
tumor-suppressor genes are present. Beyond this, it is possible
to make this statement for all genes of an organism. The
Zygosity-Predictor was primarily developed to handle SNVs and
INDELs (later addressed as small-variants) of somatic and
germline origin. In order not to overlook severe effects
outside of the small-variant context, it has been extended with
the assessment of large scale deletions, which cause losses of
whole genes or parts of them." --Marco Rheinnecker, Marc Ruebsam, Daniel Huebschmann, Martina Froehlich, Barbara Hutter\\
CancerInSilico & "The CancerInSilico package provides an R interface for
running mathematical models of tumor progresson and generating
gene expression data from the results. This package has the
underlying models implemented in C++ and the output and
analysis features implemented in R." --Thomas D. Sherman, Raymond Cheng, Elana J. Fertig\\
\addlinespace
CancerSubtypes & "CancerSubtypes integrates the current common computational
biology methods for cancer subtypes identification and provides
a standardized framework for cancer subtype analysis based
multi-omics data, such as gene expression, miRNA expression,
DNA methylation and others." --Taosheng Xu\\
IRISFGM & "Single-cell RNA-Seq data is useful in discovering cell
heterogeneity and signature genes in specific cell populations
in cancer and other complex diseases. Specifically, the
investigation of functional gene modules (FGM) can help to
understand gene interactive networks and complex biological
processes. QUBIC2 is recognized as one of the most efficient
and effective tools for FGM identification from scRNA-Seq data.
However, its availability is limited to a C implementation, and
its applicative power is affected by only a few downstream
analyses functionalities. We developed an R package named
IRIS-FGM (integrative scRNA-Seq interpretation system for
functional gene module analysis) to support the investigation
of FGMs and cell clustering using scRNA-Seq data. Empowered by
QUBIC2, IRIS-FGM can identify co-expressed and co-regulated
FGMs, predict types/clusters, identify differentially expressed
genes, and perform functional enrichment analysis. It is
noteworthy that IRIS-FGM also applies Seurat objects that can
be easily used in the Seurat vignettes." --Yuzhou Chang, Qin Ma, Carter Allen, Dongjun Chung\\
STROMA4 & "This package estimates four stromal properties identified
in TNBC patients in each patient of a gene expression datasets.
These stromal property assignments can be combined to subtype
patients. These four stromal properties were identified in
Triple negative breast cancer (TNBC) patients and represent the
presence of different cells in the stroma: T-cells (T), B-cells
(B), stromal infiltrating epithelial cells (E), and desmoplasia
(D). Additionally this package can also be used to estimate
generative properties for the Lehmann subtypes, an alternative
TNBC subtyping scheme (PMID: 21633166)." --Sadiq Saleh, Michael Hallett\\
HPAStainR & "This package is built around the HPAStainR function. The
purpose of the HPAStainR function is to query the visual
staining data in the Human Protein Atlas to return a table of
staining ranked cell types. The function also has multiple
arguments to personalize to output as well to include cancer
data, csv readable names, modify the confidence levels of the
results and more. The other functions exist exclusively to
easily acquire the data required to run HPAStainR." --Tim O. Nieuwenhuis\\
\bottomrule
\end{longtable}

\newpage

\section{Bioconductor data packages with `cancer' in package description}\label{app2}

Quoted text is the content of the Description element of the package DESCRIPTION. Names following the
text are as reported in the Authors field.

\begin{longtable}[t]{l>{\raggedright\arraybackslash}p{25em}}
\toprule
package & desc\\
\midrule
antiProfilesData & "Colon normal tissue and cancer samples used in Corrada
Bravo, et al. gene expression anti-profiles paper: BMC
Bioinformatics 2012, 13:272 doi:10.1186/1471-2105-13-272.
Measurements are z-scores obtained from the GeneExpression
Barcode in the 'frma' package" --Hector Corrada Bravo, Matthew McCall, Rafael A. Irizarry\\
BloodCancerMultiOmics2017 & "The package contains data of the Primary Blood Cancer
Encyclopedia (PACE) project together with a complete executable
transcript of the statistical analysis and reproduces figures
presented in the paper "Drug-perturbation-based stratification
of blood cancer" by Dietrich S, Oles M, Lu J et al., J. Clin.
Invest. (2018) 128(1):427-445. doi:10.1172/JCI93801." --Malgorzata Oles, Sascha Dietrich, Junyan Lu, Britta Velten, Andreas Mock, Vladislav Kim, Wolfgang Huber\\
breastCancerMAINZ & "Gene expression data from the breast cancer study
published by Schmidt et al. in 2008, provided as an eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
breastCancerNKI & "Genexpression data from a breast cancer study published by
van't Veer et al. in 2002 and van de Vijver et al. in 2002,
provided as an eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
breastCancerTRANSBIG & "Gene expression data from a breast cancer study published
by Desmedt et al. in 2007, provided as an eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
\addlinespace
breastCancerUNT & "Gene expression data from a breast cancer study published
by Sotiriou et al. in 2007, provided as an eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
breastCancerUPP & "Gene expression data from a breast cancer study published
by Miller et al. in 2005, provided as an eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
breastCancerVDX & "Gene expression data from a breast cancer study published
by Wang et al. in 2005 and Minn et al. in 2007, provided as an
eSet." --Markus Schroeder, Benjamin Haibe-Kains, Aedin Culhane, Christos Sotiriou, Gianluca Bontempi, John Quackenbush\\
cancerdata & "Dataset for the R package cancerclass" --Jan Budczies, Daniel Kosztyla\\
cfToolsData & "The cfToolsData package supplies the data for the cfTools
package. It contains two pre-trained deep neural network (DNN)
models for the cfSort function. Additionally, it includes the
shape parameters of beta distribution characterizing
methylation markers associated with four tumor types for the
CancerDetector function, as well as the parameters
characterizing methylation markers specific to 29 primary human
tissue types for the cfDeconvolve function." --Ran Hu, Shuo Li, Xianghong Jasmine Zhou, Wenyuan Li\\
\addlinespace
CLLmethylation & "The package includes DNA methylation data for the primary
Chronic Lymphocytic Leukemia samples included in the Primary
Blood Cancer Encyclopedia (PACE) project. Raw data from the
450k DNA methylation arrays is stored in the European
Genome-Phenome Archive (EGA) under accession number
EGAS0000100174. For more information concerning the project
please refer to the paper "Drug-perturbation-based
stratification of blood cancer" by Dietrich S, Oles M, Lu J et
al., J. Clin. Invest. (2018) and R/Bioconductor package
BloodCancerMultiOmics2017." --Malgorzata Oles, Andreas Mock\\
colonCA & "exprSet for Alon et al. (1999) colon cancer data" --Sylvia Merk\\
COSMIC.67 & "COSMIC: Catalogue Of Somatic Mutations In Cancer, version
67 (2013-10-24)" --Julian Gehring\\
CRCL18 & "colorectal cancer mRNA and miRNA on 18 cell lines" --Claudio Isella\\
curatedBladderData & "The curatedBladderData package provides relevant functions
and data for gene expression analysis in patients with bladder
cancer." --Markus Riester\\
\addlinespace
curatedBreastData & "Curated human breast cancer tissue S4 ExpresionSet
datasets from over 16 clinical trials comprising over 2,000
patients. All datasets contain at least one type of outcomes
variable and treatment information (minimum level: whether they
had chemotherapy and whether they had hormonal therapy).
Includes code to post-process these datasets." --Katie Planey\\
curatedCRCData & "The curatedCRC package provides relevant functions and
data for gene expression analysis in patients with colorectal
cancer." --Princy Parsana, Markus Riester, Curtis Huttenhower, Levi Waldron\\
curatedOvarianData & "The curatedOvarianData package provides data for gene
expression analysis in patients with ovarian cancer." --Benjamin F. Ganzfried, Markus Riester, Steve Skates, Victoria Wang, Thomas Risch, Benjamin Haibe-Kains, Svitlana Tyekucheva, Jie Ding, Ina Jazic, Michael Birrer, Giovanni Parmigiani, Curtis Huttenhower, Levi Waldron\\
curatedTCGAData & "This package provides publicly available data from The
Cancer Genome Atlas (TCGA) as MultiAssayExperiment objects.
MultiAssayExperiment integrates multiple assays (e.g., RNA-seq,
copy number, mutation, microRNA, protein, and others) with
clinical / pathological data. It also links assay barcodes with
patient identifiers, enabling harmonized subsetting of rows
(features) and columns (patients / samples) across the entire
multi-'omics experiment." --Marcel Ramos, Levi Waldron, Lucas Schiffer, Ludwig Geistlinger, Valerie Obenchain, Martin Morgan\\
depmap & "The depmap package is a data package that accesses datsets
from the Broad Institute DepMap cancer dependency study using
ExperimentHub. Datasets from the most current release are
available, including RNAI and CRISPR-Cas9 gene knockout screens
quantifying the genetic dependency for select cancer cell
lines. Additional datasets are also available pertaining to the
log copy number of genes for select cell lines, protein
expression of cell lines as measured by reverse phase protein
lysate microarray (RPPA), 'Transcript Per Million' (TPM) data,
as well as supplementary datasets which contain metadata and
mutation calls for the other datasets found in the current
release. The 19Q3 release adds the drug\_dependency dataset,
that contains cancer cell line dependency data with respect to
drug and drug-candidate compounds. The 20Q2 release adds the
proteomic dataset that contains quantitative profiling of
proteins via mass spectrometry. This package will be updated on
a quarterly basis to incorporate the latest Broad Institute
DepMap Public cancer dependency datasets. All data made
available in this package was generated by the Broad Institute
DepMap for research purposes and not intended for clinical use.
This data is distributed under the Creative Commons license
(Attribution 4.0 International (CC BY 4.0))." --Laurent Gatto, Theo Killian\\
\addlinespace
easierData & "Access to internal data required for the functional
performance of easier package and exemplary bladder cancer
dataset with both processed RNA-seq data and information on
response to ICB therapy generated by Mariathasan et al. "TGF-B
attenuates tumour response to PD-L1 blockade by contributing to
exclusion of T cells", published in Nature, 2018
[doi:10.1038/nature25501](https://doi.org/10.1038/nature25501).
The data is made available via
[`IMvigor210CoreBiologies`](http://research-pub.gene.com/IMvigor210CoreBiologies/)
package under the CC-BY license." --Oscar Lapuente-Santana, Federico Marini, Arsenij Ustjanzew, Francesca Finotello, Federica Eduati\\
fabiaData & "Supplying gene expression data sets for the demos of the
biclustering method "Factor Analysis for Bicluster Acquisition"
(FABIA). The following three data sets are provided: A) breast
cancer (van't Veer, Nature, 2002), B) multiple tissues (Su,
PNAS, 2002), and C) diffuse large-B-cell lymphoma (Rosenwald, N
Engl J Med, 2002)." --Sepp Hochreiter\\
gageData & "This is a supportive data package for the software
package, gage. However, the data supplied here are also useful
for gene set or pathway analysis or microarray data analysis in
general. In this package, we provide two demo microarray
dataset: GSE16873 (a breast cancer dataset from GEO) and BMP6
(originally published as an demo dataset for GAGE, also
registered as GSE13604 in GEO). This package also includes
commonly used gene set data based on KEGG pathways and GO terms
for major research species, including human, mouse, rat and
budding yeast. Mapping data between common gene IDs for budding
yeast are also included." --Weijun Luo\\
GSE62944 & "TCGA processed RNA-Seq data for 9264 tumor and 741 normal
samples across 24 cancer types and made them available as GEO
accession
[GSE62944](http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE62944).
GSE62944 data have been parsed into a SummarizedExperiment
object available in ExperimentHub." --Sonali Arora\\
GSVAdata & "This package stores the data employed in the vignette of
the GSVA package. These data belong to the following
publications: Armstrong et al. Nat Genet 30:41-47, 2002; Cahoy
et al. J Neurosci 28:264-278, 2008; Carrel and Willard, Nature,
434:400-404, 2005; Huang et al. PNAS, 104:9758-9763, 2007;
Pickrell et al. Nature, 464:768-722, 2010; Skaletsky et al.
Nature, 423:825-837; Verhaak et al. Cancer Cell 17:98-110, 2010" --Robert Castelo\\
\addlinespace
HarmonizedTCGAData & "This package contains the processed harmonized TCGA data
of five cancer types used in "Tianle Ma and Aidong Zhang,
Integrate Multi-omic Data Using Affinity Network Fusion (ANF)
for Cancer Patient Clustering"." --Tianle Ma\\
HD2013SGI & "This package contains the experimental data and a complete
executable transcript (vignette) of the analysis of the HCT116
genetic interaction matrix presented in the paper "Mapping
genetic interactions in human cancer cells with RNAi and
multiparametric phenotyping" by C. Laufer, B. Fischer, M.
Billmann, W. Huber, M. Boutros; Nature Methods (2013)
10:427-31. doi: 10.1038/nmeth.2436." --Bernd Fischer\\
LungCancerACvsSCCGEO & "This package contains 30 Affymetrix CEL files for 7
Adenocarcinoma (AC) and 8 Squamous cell carcinoma (SCC) lung
cancer samples taken at random from 3 GEO datasets (GSE10245,
GSE18842 and GSE2109) and other 15 samples from a dataset
produced by the organizers of the IMPROVER Diagnostic Signature
Challenge available from GEO (GSE43580)." --Adi Laurentiu Tarca\\
LungCancerLines & "Reads from an RNA-seq experiment between two lung cancer
cell lines: H1993 (met) and H2073 (primary). The reads are
stored as Fastq files and are meant for use with the TP53Genome
object in the gmapR package." --Cory Barr, Michael Lawrence\\
lungExpression & "Data from three large lung cancer studies provided as
ExpressionSets" --Robert Scharpf, Simens Zhong, Giovanni Parmigiani\\
\addlinespace
mammaPrintData & "Gene expression data for the two breast cancer cohorts
published by Glas and Buyse in 2006. This cohorts were used to
implement and validate the mammaPrint breast cancer test." --Luigi Marchionni\\
mAPKLData & "Gene expression data from a breast cancer study published
by Turashvili et al. in 2007, provided as an eSet." --Argiris Sakellariou\\
mcsurvdata & "This package stores two merged expressionSet objects that
contain the gene expression profile and clinical information of
-a- six breast cancer cohorts and -b- four colorectal cancer
cohorts. Breast cancer data are employed in the vignette of the
hrunbiased package for survival analysis of gene signatures." --Adria Caballe Mestres, Antoni Berenguer Llergo, Camille Stephan-Otto Attolini\\
MetaGxBreast & "A collection of Breast Cancer Transcriptomic Datasets that
are part of the MetaGxData package compendium." --Michael Zon, Deena M.A. Gendoo, Christopher Eeles, Benjamin Haibe-Kains\\
MetaGxOvarian & "A collection of Ovarian Cancer Transcriptomic Datasets
that are part of the MetaGxData package compendium." --Michael Zon, Vandana Sandhu, Christopher Eeles, Benjamin Haibe-Kains\\
\addlinespace
MetaGxPancreas & "A collection of pancreatic Cancer transcriptomic datasets
that are part of the MetaGxData package compendium. This
package contains multiple pancreas cancer datasets that have
been downloaded from various resources and turned into
SummarizedExperiment objects. The details of how the authors
normalized the data can be found in the experiment data section
of the objects. Additionally, the location the data was
obtained from can be found in the url variables of the
experiment data portion of each SE." --Michael Zon, Vandana Sandhu, Christopher Eeles, Benjamin Haibe-Kains\\
microRNAome & "This package provides a SummarizedExperiment object of
read counts for microRNAs across tissues, cell-types, and
cancer cell-lines. The read count matrix was prepared and
provided by the author of the study: Towards the human cellular
microRNAome." --Matthew N. McCall, Marc K. Halushka, Arun H. Patil\\
NGScopyData & "Subset of BAM files of human lung tumor and pooled normal
samples by targeted panel sequencing. [Zhao et al 2014.
Targeted Sequencing in Non-Small Cell Lung Cancer (NSCLC) Using
the University of North Carolina (UNC) Sequencing Assay
Captures Most Previously Described Genetic Aberrations in
NSCLC. In preparation.] Each sample is a 10 percent random
subsample drawn from the original sequencing data. The pooled
normal sample has been rescaled accroding to the total number
of normal samples in the "pool". Here provided is the
subsampled data on chr6 (hg19)." --Xiaobei Zhao\\
octad.db & "Open Cancer TherApeutic Discovery (OCTAD) package implies
sRGES approach for the drug discovery. The essential idea is to
identify drugs that reverse the gene expression signature of a
disease by tamping down over-expressed genes and stimulating
weakly expressed ones. The following package contains all
required precomputed data for whole OCTAD pipeline computation." --E. Chekalin, S. Paithankar, B. Zeng, B. Glicksberg, P. Newbury, J. Xing, K. Liu, A. Wen, D. Joseph, B. Chen\\
ProData & "A data package of SELDI-TOF protein mass spectrometry data
of 167 breast cancer and normal samples." --Xiaochun Li\\
\addlinespace
prostateCancerCamcap & "A Bioconductor data package for the Ross-Adams (2015)
Prostate Cancer dataset." --Mark Dunning\\
prostateCancerGrasso & "A Bioconductor data package for the Grasso (2012) Prostate
Cancer dataset." --Mark Dunning\\
rcellminerData & "The NCI-60 cancer cell line panel has been used over the
course of several decades as an anti-cancer drug screen. This
panel was developed as part of the Developmental Therapeutics
Program (DTP, http://dtp.nci.nih.gov/) of the U.S. National
Cancer Institute (NCI). Thousands of compounds have been tested
on the NCI-60, which have been extensively characterized by
many platforms for gene and protein expression, copy number,
mutation, and others (Reinhold, et al., 2012). The purpose of
the CellMiner project (http://discover.nci.nih.gov/ cellminer)
has been to integrate data from multiple platforms used to
analyze the NCI-60 and to provide a powerful suite of tools for
exploration of NCI-60 data." --Augustin Luna, Vinodh Rajapakse, Fabricio Sousa\\
RTCGA.clinical & "Package provides clinical datasets from The Cancer Genome
Atlas Project for all cohorts types from
http://gdac.broadinstitute.org/. Clinical data format is
explained here
https://wiki.nci.nih.gov/display/TCGA/Clinical+Data+Overview.
Data from 2015-11-01 snapshot." --Marcin Kosinski\\
RTCGA.CNV & "Package provides CNV (based on Merge snp) datasets from
The Cancer Genome Atlas Project for all cohorts types from
http://gdac.broadinstitute.org/. Data format is explained here
https://wiki.nci.nih.gov/display/TCGA/Retrieving
+Data+Using+the+Data+Matrix. Data from 2015-11-01 snapshot." --Przemyslaw Biecek\\
\addlinespace
RTCGA.methylation & "Package provides methylation (humanmethylation27) datasets
from The Cancer Genome Atlas Project for all available cohorts
types from http://gdac.broadinstitute.org/. Data format is
explained here
https://wiki.nci.nih.gov/display/TCGA/DNA+methylation Data from
2015-11-01 snapshot." --Marcin Kosinski, Witold Chodor\\
RTCGA.miRNASeq & "Package provides miRNASeq datasets from The Cancer Genome
Atlas Project for all available cohorts types from
http://gdac.broadinstitute.org/. Data format is explained here
https://wiki.nci.nih.gov/display/TCGA/miRNASeq\#miRNASeq-DataOverview
Data from 2015-11-01 snapshot." --Witold \vphantom{1} Chodor\\
RTCGA.mRNA & "Package provides mRNA datasets from The Cancer Genome
Atlas Project for all available cohorts types from
http://gdac.broadinstitute.org/. Data format is explained here
https://wiki.nci.nih.gov/display /TCGA/Gene+expression+data Data
from 2015-11-01 snapshot." --Witold Chodor\\
RTCGA.mutations & "Package provides mRNA datasets from The Cancer Genome
Atlas Project for all available cohorts types from
http://gdac.broadinstitute.org/. Data format is explained here
https://wiki.nci.nih.gov/display /TCGA/Gene+expression+data Data
from 2015-11-01 snapshot." --Witold Chodor\\
RTCGA.PANCAN12 & "Package provides clinical, expression, cnv and mutation
data from Genome Cancer Browser." --Przemyslaw Biecek\\
\addlinespace
RTCGA.rnaseq & "Package provides rna-seq datasets from The Cancer Genome
Atlas Project for all cohorts types from
http://gdac.broadinstitute.org/. Rna-seq data format is
explained here
https://wiki.nci.nih.gov/display/TCGA/RNASeq+Version+2. Data
source is illumina hiseq Level 3 RSEM normalized expression
data. Data from 2015-11-01 snapshot." --Marcin Kosinski\\
RTCGA.RPPA & "Package provides RPPA datasets from The Cancer Genome
Atlas Project for all available cohorts types from
http://gdac.broadinstitute.org/. Data format is explained here
https://wiki.nci.nih.gov/display/TCGA/Protein+Array
+Data+Format+Specification?src=search" --Witold Chodor\\
SBGNview.data & "This package contains: 1. A microarray gene expression
dataset from a human breast cancer study. 2. A RNA-Seq gene
expression dataset from a mouse study on IFNG knockout. 3. ID
mapping tables between gene IDs and SBGN-ML file glyph IDs. 4.
Percent of orthologs detected in other species of the genes in
a pathway. Cutoffs of this percentage for defining if a pathway
exists in another species. 5. XML text of SBGN-ML files for all
pre-collected pathways." --Xiaoxi Dong*, Kovidh Vegesna*, Weijun Luo\\
seventyGeneData & "Gene expression data for the two breast cancer cohorts
published by van't Veer and Van de Vijver in 2002." --Luigi Marchionni, Claudio Zanettini, Lucio Queiroz\\
SFEData & "Example spatial transcriptomics datasets with Simple
Feature annotations as SpatialFeatureExperiment objects.
Technologies include Visium, slide-seq, Nanostring CoxMX,
Vizgen MERFISH, and 10X Xenium. Tissues include mouse skeletal
muscle, human melanoma metastasis, human lung, breast cancer,
and mouse liver." --Lambda Moses, Kayla Jackson, Lior Pachter\\
\addlinespace
shinyMethylData & "Extracted data from 369 TCGA Head and Neck Cancer DNA
methylation samples. The extracted data serve as an example
dataset for the package shinyMethyl. Original samples are from
450k methylation arrays, and were obtained from The Cancer
Genome Atlas (TCGA). 310 samples are from tumor, 50 are matched
normals and 9 are technical replicates of a control cell line." --Jean-Philippe Fortin, Kasper Daniel Hansen\\
SomatiCAData & "An example cancer whole genome sequencing data for the
SomatiCA package" --Mengjie Chen\\
SomaticCancerAlterations & "Collection of somatic cancer alteration datasets" --Julian Gehring\\
stjudem & "This is a microarray data set on acute lymphoblastic
leukemia, published in 2002 (Yeoh et al.Cancer Cell 2002). The
experiments were conducted in the St.Jude Children's Research
Hospital, Memphis, Tenessee, USA. The raw data was preprocessed
by variance stabilizing normalization (Huber et al.) on probe
and subsequent summarization of probe expression values into
probe set expression values using median polish." --Benjamin Georgi, Matthias Heinig, Sebastian Schmeier, Joern Toedling\\
TCGAcrcmiRNA & "colorectal cancer miRNA profile provided by TCGA" --Claudio Isella\\
\addlinespace
TCGAcrcmRNA & "colorectal cancer mRNA profile provided by TCGA" --Claudio Isella\\
TCGAMethylation450k & "The Cancer Genome Atlas (TCGA) is applying genomics
technologies to over 20 different types of cancer.  This
package contains a small set of 450k array data in idat format." --Sean Davis\\
TCGAWorkflowData & "This experimental data package contains 11 data sets
necessary to follow the "TCGA Workflow: Analyze cancer genomics
and epigenomics data using Bioconductor packages"." --Tiago Chedraoui Silva\\
\bottomrule
\end{longtable}

