\section*{References}\label{refs}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Aevermann2018}{}%
Aevermann, Brian D., Mark Novotny, Trygve Bakken, Jeremy A. Miller, Alexander D. Diehl, David Osumi-Sutherland, Roger S. Lasken, Ed S. Lein, and Richard H. Scheuermann. 2018. ``Cell Type Discovery Using Single-Cell Transcriptomics: Implications for Ontological Representation.'' \emph{Human Molecular Genetics} 27 (R1): R40--R47. \url{https://doi.org/10.1093/hmg/ddy100}.

\leavevmode\hypertarget{ref-Aganezov2022}{}%
Aganezov, Sergey, Stephanie M. Yan, Daniela C. Soto, Melanie Kirsche, Samantha Zarate, Pavel Avdeyev, Dylan J. Taylor, et al. 2022. ``A Complete Reference Genome Improves Analysis of Human Genetic Variation.'' \emph{Science} 376 (6588). \url{https://doi.org/10.1126/science.abl3533}.

\leavevmode\hypertarget{ref-Aibar2015}{}%
Aibar, Sara, Celia Fontanillo, Conrad Droste, and Javier De Las Rivas. 2015. ``Functional Gene Networks: R/Bioc Package to Generate and Analyse Gene Networks Derived from Functional Enrichment and Clustering.'' \emph{Bioinformatics} 31 (10): 1686--8. \url{https://doi.org/10.1093/bioinformatics/btu864}.

\leavevmode\hypertarget{ref-Argelaguet2020}{}%
Argelaguet, Ricard, Damien Arnol, Danila Bredikhin, Yonatan Deloro, Britta Velten, John C. Marioni, and Oliver Stegle. 2020. ``MOFA+: A Statistical Framework for Comprehensive Integration of Multi-Modal Single-Cell Data.'' \emph{Genome Biology} 21 (1). \url{https://doi.org/10.1186/s13059-020-02015-1}.

\leavevmode\hypertarget{ref-Aryee2014}{}%
Aryee, Martin J., Andrew E. Jaffe, Hector Corrada-Bravo, Christine Ladd-Acosta, Andrew P. Feinberg, Kasper D. Hansen, and Rafael A. Irizarry. 2014. ``Minfi: A Flexible and Comprehensive Bioconductor Package for the Analysis of Infinium Dna Methylation Microarrays.'' \emph{Bioinformatics} 30 (10): 1363--9. \url{https://doi.org/10.1093/bioinformatics/btu049}.

\leavevmode\hypertarget{ref-Cavalcante2017}{}%
Cavalcante, Raymond G, and Maureen A Sartor. 2017. ``Annotatr: Genomic Regions in Context.'' Edited by Alfonso Valencia. \emph{Bioinformatics} 33 (15): 2381--3. \url{https://doi.org/10.1093/bioinformatics/btx183}.

\leavevmode\hypertarget{ref-Couto2023}{}%
Couto, Bárbara Zita Peters, Nicholas Robertson, Ellis Patrick, and Shila Ghazanfar. 2023. ``MoleculeExperiment Enables Consistent Infrastructure for Molecule-Resolved Spatial Transcriptomics Data in Bioconductor.'' \emph{bioRxiv}. \url{https://doi.org/10.1101/2023.05.16.541040}.

\leavevmode\hypertarget{ref-Gel2015}{}%
Gel, Bernat, Anna Diez-Villanueva, Eduard Serra, Marcus Buschbeck, Miguel A. Peinado, and Roberto Malinverni. 2015. ``regioneR: An R/Bioconductor Package for the Association Analysis of Genomic Regions Based on Permutation Tests.'' \emph{Bioinformatics}, September, btv562. \url{https://doi.org/10.1093/bioinformatics/btv562}.

\leavevmode\hypertarget{ref-Hansen2012}{}%
Hansen, Kasper D., Rafael a. Irizarry, and Zhijin Wu. 2012. ``Removing Technical Variability in Rna-Seq Data Using Conditional Quantile Normalization.'' \emph{Biostatistics} 13: 204--16. \url{https://doi.org/10.1093/biostatistics/kxr054}.

\leavevmode\hypertarget{ref-ICBnci}{}%
``Immune Checkpoint Inhibitors.'' 2022. \url{https://www.cancer.gov/about-cancer/treatment/types/immunotherapy/checkpoint-inhibitors}.

\leavevmode\hypertarget{ref-Kupkova2022}{}%
Kupkova, Kristyna, Jose Verdezoto Mosquera, Jason P. Smith, Michał Stolarczyk, Tessa L. Danehy, John T. Lawson, Bingjie Xue, John T. Stubbs, Nathan LeRoy, and Nathan C. Sheffield. 2022. ``GenomicDistributions: Fast Analysis of Genomic Intervals with Bioconductor.'' \emph{BMC Genomics} 23 (1). \url{https://doi.org/10.1186/s12864-022-08467-y}.

\leavevmode\hypertarget{ref-easierPap}{}%
Lapuente-Santana, Óscar, Maisa van Genderen, Peter A. J. Hilbers, Francesca Finotello, and Federica Eduati. 2021. ``Interpretable Systems Biomarkers Predict Response to Immune-Checkpoint Inhibitors.'' \emph{Patterns} 2 (8). \url{https://doi.org/10.1016/j.patter.2021.100293}.

\leavevmode\hypertarget{ref-Lawson2018}{}%
Lawson, John, Eleni Tomazou, Christoph Bock, and Nathan C. Sheffield. 2018. ``MIRA: An R Package for DNA Methylation-Based Inference of Regulatory Activity.'' \emph{Bioinformatics} bty083 (March). \url{https://doi.org/10.1093/bioinformatics/bty083}.

\leavevmode\hypertarget{ref-Lawson2020}{}%
Lawson, John T., Jason P. Smith, Stefan Bekiranov, Francine E. Garrett-Bakelman, and Nathan C. Sheffield. 2020. ``COCOA: Coordinate Covariation Analysis of Epigenetic Heterogeneity.'' \emph{Genome Biology} 21 (1). \url{https://doi.org/10.1186/s13059-020-02139-4}.

\leavevmode\hypertarget{ref-Mariathasan2018}{}%
Mariathasan, Sanjeev, Shannon J. Turley, Dorothee Nickles, Alessandra Castiglioni, Kobe Yuen, Yulei Wang, Edward E. Kadel III, et al. 2018. ``TGFB Attenuates Tumour Response to Pd-L1 Blockade by Contributing to Exclusion of T Cells.'' \emph{Nature} 554 (7693): 544--48. \url{https://doi.org/10.1038/nature25501}.

\leavevmode\hypertarget{ref-moses23}{}%
Moses, Lambda, Pétur Helgi Einarsson, Kayla Jackson, Laura Luebbert, A Sina Booeshaghi, Sindri Antonsson, Nicolas Bray, Páll Melsted, and Lior Pachter. 2023. ``Voyager: Exploratory Single-Cell Genomics Data Analysis with Geospatial Statistics.'' \emph{bioRxiv}.

\leavevmode\hypertarget{ref-Mueller2019}{}%
Müller, Fabian, Michael Scherer, Yassen Assenov, Pavlo Lutsik, Jörn Walter, Thomas Lengauer, and Christoph Bock. 2019. ``RnBeads 2.0: Comprehensive Analysis of Dna Methylation Data.'' \emph{Genome Biology} 20 (1). \url{https://doi.org/10.1186/s13059-019-1664-9}.

\leavevmode\hypertarget{ref-Ou2018}{}%
Ou, Jianhong, Haibo Liu, Jun Yu, Michelle A. Kelliher, Lucio H. Castilla, Nathan D. Lawson, and Lihua Julie Zhu. 2018. ``ATACseqQC: A Bioconductor Package for Post-Alignment Quality Assessment of Atac-Seq Data.'' \emph{BMC Genomics} 19 (1). \url{https://doi.org/10.1186/s12864-018-4559-3}.

\leavevmode\hypertarget{ref-rig22}{}%
Righelli, Dario, Lukas M Weber, Helena L Crowell, Brenda Pardo, Leonardo Collado-Torres, Shila Ghazanfar, Aaron TL Lun, Stephanie C Hicks, and Davide Risso. 2022. ``SpatialExperiment: Infrastructure for Spatially-Resolved Transcriptomics Data in R Using Bioconductor.'' \emph{Bioinformatics} 38 (11): 3128--31.

\leavevmode\hypertarget{ref-Schatz2022}{}%
Schatz, Michael C., Anthony A. Philippakis, Enis Afgan, Eric Banks, Vincent J. Carey, Robert J. Carroll, Alessandro Culotti, et al. 2022. ``Inverting the Model of Genomics Data Sharing with the Nhgri Genomic Data Science Analysis, Visualization, and Informatics Lab-Space.'' \emph{Cell Genomics} 2 (1): 100085. \url{https://doi.org/10.1016/j.xgen.2021.100085}.

\leavevmode\hypertarget{ref-Sheffield2016}{}%
Sheffield, Nathan C., and Christoph Bock. 2016. ``LOLA: Enrichment Analysis for Genomic Region Sets and Regulatory Elements in R and Bioconductor.'' \emph{Bioinformatics} 32 (4): 587--89. \url{https://doi.org/10.1093/bioinformatics/btv612}.

\leavevmode\hypertarget{ref-Silva2019}{}%
Silva, Tiago C., Simon G. Coetzee, Nicole Gull, Lijing Yao, Dennis J. Hazelett, Houtan Noushmehr, De-Chen Lin, and Benjamin P. Berman. 2018. ``ELMER V.2: An R/Bioconductor Package to Reconstruct Gene Regulatory Networks from Dna Methylation and Transcriptome Profiles.'' Edited by Oliver Stegle. \emph{Bioinformatics} 35 (11): 1974--7. \url{https://doi.org/10.1093/bioinformatics/bty902}.

\leavevmode\hypertarget{ref-Stark2011}{}%
Stark, Rory, and Gordon Brown. 2011. ``DiffBind: Differential Binding Analysis of Chip-Seq Peak Data.''

\leavevmode\hypertarget{ref-Tian2018}{}%
Tian, Luyi, Shian Su, Xueyi Dong, Daniela Amann-Zalcenstein, Christine Biben, Azadeh Seidi, Douglas J. Hilton, Shalin H. Naik, and Matthew E. Ritchie. 2018. ``ScPipe: A Flexible R/Bioconductor Preprocessing Pipeline for Single-Cell Rna-Sequencing Data.'' Edited by Mihaela Pertea. \emph{PLOS Computational Biology} 14 (8): e1006361. \url{https://doi.org/10.1371/journal.pcbi.1006361}.

\leavevmode\hypertarget{ref-webr23}{}%
Weber, Lukas M, Arkajyoti Saha, Abhirup Datta, Kasper D Hansen, and Stephanie C Hicks. 2023. ``NnSVG for the Scalable Identification of Spatially Variable Genes Using Nearest-Neighbor Gaussian Processes.'' \emph{Nature Communications} 14 (1): 4059.

\leavevmode\hypertarget{ref-Wei2018}{}%
Wei, Zheng, Wei Zhang, Huan Fang, Yanda Li, and Xiaowo Wang. 2018. ``EsATAC: An Easy-to-Use Systematic Pipeline for Atac-Seq Data Analysis.'' \emph{Bioinformatics (Oxford, England)}, March. \url{https://doi.org/10.1093/bioinformatics/bty141}.

\leavevmode\hypertarget{ref-Zheng2018a}{}%
Zheng, Shijie C., Charles E. Breeze, Stephan Beck, and Andrew E. Teschendorff. 2018. ``Identification of Differentially Methylated Cell Types in Epigenome-Wide Association Studies.'' \emph{Nature Methods} 15 (12): 1059--66. \url{https://doi.org/10.1038/s41592-018-0213-x}.

\leavevmode\hypertarget{ref-Zhu2010}{}%
Zhu, Lihua J, Claude Gazin, Nathan D Lawson, Hervé Pagès, Simon M Lin, David S Lapointe, and Michael R Green. 2010. ``ChIPpeakAnno: A Bioconductor Package to Annotate ChIP-Seq and ChIP-Chip Data.'' \emph{BMC Bioinformatics} 11 (1). \url{https://doi.org/10.1186/1471-2105-11-237}.
\end{cslreferences}


\end{document}
